---
title: "Guide to WOODCARB3R"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Guide to WOODCARB3R}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---
```{r,echo=FALSE, message=FALSE}
knitr::opts_chunk$set(collapse = T, comment = NA)
library(WOODCARB3R)
```

This package is designed to calculate various statistics regarding wood products in the United States. Included are functions to calculate carbon from dumps/landfills, paper and wood products of different wood types. Statistics are broken down into production, imports and exports. 

The main goal is to calculate the final carbon contribution which is reported to the IPCC. This is calculated based on the approach selected (Production, Stock Change, Atmospheric).  

Currently, the Production Approach is the one used to report contribution values to the IPCC. 

###Final Carbon Contribution  

The final reported value is the contribution to AFOLU CO<sub>2</sub> emissions/removal. It's units are Gg CO<sub>2</sub>/Yr 

To calculate the carbon contribution: 

* `finalCarbonContribution()` 

The default arguments calculate the carbon contribution from 1990 to 2015 using the Production approach. It returns a single vector with the carbon contributions for each year specified. 

```{r}
finalCarbonContribution()
```

Different ranges of years can be selected along with a seperate approach. 

```{r}
finalCarbonContribution(Years = 1945:2017, approach = "Stock Change")
```

Other options include calculating using a different decay type or values for half-lifes. 

```{r}
#finalCarbonContribution(decaytype = "K=2", halflives = halfLives * 1.2)
```

A plot can be returned as well. 

```{r, fig.width=6, fig.height=4}
finalCarbonContribution(plot = TRUE)
```

###Individual Variables  

The final carbon contribution from `finalCarbonContribution()` is calculate from several variables. 

`finalVariables()` can be used to calculate individual variables. 

The variables under each approach are used to calculate that approach only. 

####Production Approach  

`2A` - Annual Change in stock of HWP in use produced from domestic harvest ($\Delta C_{HWP \: IU \ DH}$) 

`2B` - Annual Change in stock of HWP in SWDS produced from domestic harvest ($\Delta C_{HWP \: 
SWDS \: DH}$)
```{r}
finalVariables(Variable = "Var2A") 
finalVariables(Variable = "Var2B")
```

####Stock Change Approach  

`1A` - Annual Change in stock of HWP in use form consumption ($\Delta C_{HWP \: IU \: DC}$)  

`1B` - Annual Change in stock of HWP in SWDS from consumption ($\Delta C_{HWP \: SWDS \: DC}$) 

```{r}
finalVariables(Variable = "Var1A")
finalVariables(Variable = "Var1B")
```

###Other Statistics 

`finalCarbonContribution` and `finalVariables` calculate the final values for the WOODCARB model. There are many intermediate calculations and statistics that can be accessed using the package that are necessary to calculate the final values. 

####Total Carbon 

The model considers carbon stored in solid wood products, paper and solid waste disposal sites (SWDS). There are functions to calculate amounts of carbon stored in these places. 

#####Solid Wood Products 

`swpcarbontotal()` calculates total carbon stored in solid wood products in oven-dry tons of wood fiber.

It gives the option of returning carbon totals by end use: (the default argument returns total carbon for all end uses)

```{r}
swpcarbontotal(c(1990,2015), onlytotal = FALSE) 
```

The default approach is Production but carbon totals using the Stock Change approach can be calculated as well. 

```{r}
swpcarbontotal(approach = "Stock Change")
```

There is an option to exclude carbon from lumber pre-1900 (the default includes this). 

```{r}
swpcarbontotal(lumberpre = FALSE)
```

Similar to `finalCarbonContribution()`, alternate half-lives can be provided. 

```{r}
swpcarbontotal(halflives = halfLives * 1.5)
```

#####Paper  


#####Solid Waste Disposal Sites 
